### 2022-09-23 main/chapter_searches.tex
minor proofread

Standard probability algebra derives from
Equation~\ref{eqn:searches_prob_ratio}, and comprises the \textbf{product rule}
\begin{align}
\prob{c, b}{a} &= \prob{c}{b, a}\times \prob{b}{a},
\label{eqn:searches_product_rule}
\intertext{%
for nested proportions
(a spoon in a teacup in a sink),
and a \textbf{sum rule}%
}
\prob{c\;\textrm{or}\;b}{a} &= \prob{c}{a} + \prob{b}{a} - \prob{c, b}{a},
\label{eqn:searches_sum_rule}
\end{align}
for combinations in a union, with sums over exhaustive sets outcomes
necessarily normalized to unity.

@@

\subsection{\texorpdfstring{$\cls$}{CLs}}
Such misrepresentations of our results reinforces the need for clearer
replacements, which $\log$ likelihood ratios can provide.
Standard practice at the LHC contains the beginnings of such
replacements under the name $\cls$~\cite{
read2000modified,
Read2002cls,
junk1999confidence,
read1997optimal,
bock1998lower,
etde1998prospects,
lep2000searches,
lep2003search,
Murray2010heretic,
cern2011procedure,
pdg2022ynf
},
which is used for setting upper limits on disfavoured signal yields $s$
above background yields $b$:
\begin{equation}
\label{eqn:searches_cls}
\cls
=
\frac{P(s + b)}{P(b)}
=
\frac{
\prob{t \geq \tilde{t}}{s + b}
}{
\prob{t \geq \tilde{t}}{b}\hfill
}
,
\end{equation}
in which the individual \pvalues\ are named
$\clb = P(b)$ and
$\clsb = P(s + b)$.%
\footnote{%
Despite their $\mathrm{CL}$ labels, $\cls$, $\clsb$, and $\clb$ are not
Confidence Levels ---
a confidence level $\gamma$ is a threshold used to assign limits by solving
for $x$ such that $\gamma = 1 - P(x)$~\cite{pdg1996}.
In particle physics, \pvalues\ have sometimes been labelled $\mathrm{CL}$ since
at least 1984~\cite{pdg1984}, although textbooks from the '70s (which are cited
by~\cite{pdg1984}) do distinction the two~\cite{
eadie1971statistical,
frodesen1979probability
}.%s
}%
\footnote{%
Some literature defines $\clb = 1 - p_b$~\cite{
Cowan:2010js,
pdg2022ynf
}
where $p_b$ might not be a \pvalue\ if it excludes equality with the
observed test statistic~\cite{cern2011procedure}.
}

@@

How can one escape this trap?
Likelihood ratios on the actual data escape immediately by bypassing the
censor, but even inequalities need not be biased --- if taken against a
predetermined threshold, say $5$ events, then models can predict whether
$n \leq 5$, and the honestly answered truth or falsehood in data can favour
either side.
Ultimately, the reader may distrust the reporter ---
the statements
`logically, $n \leq 5$' and
`\atlas\ reports $n \leq 5$' and  are not
identical, since one might infer from the latter that in fact the original
data were $n = 5$ before their reduction into a test statistic tail.
The message itself is data, and what it says may not be the whole truth.

>>>>

Standard probability algebra derives from
Equation~\ref{eqn:searches_prob_ratio}; it comprises a \textbf{product rule}
\begin{align}
\prob{c, b}{a} &= \prob{c}{b, a}\times \prob{b}{a},
\label{eqn:searches_product_rule}
\intertext{%
for nested proportions
(a spoon in a teacup in a sink),
and a \textbf{sum rule}%
}
\prob{c\;\textrm{or}\;b}{a} &= \prob{c}{a} + \prob{b}{a} - \prob{c, b}{a},
\label{eqn:searches_sum_rule}
\end{align}
for combinations in a union, with sums over exhaustive sets outcomes
necessarily normalized to unity.

@@

\subsection{\texorpdfstring{$\cls$}{CLs}}
Such misrepresentations of our results reinforces the need for clearer
replacements, which $\log$ likelihood ratios can provide.
Standard practice at the LHC contains the beginnings of such
replacements under the name $\cls$~\cite{
read2000modified,
Read2002cls,
junk1999confidence,
read1997optimal,
bock1998lower,
etde1998prospects,
lep2000searches,
lep2003search,
Murray2010heretic,
cern2011procedure,
pdg2022ynf
},
which is used for setting upper limits on disfavoured signal yields $s$
above background yields $b$:
\begin{equation}
\label{eqn:searches_cls}
\cls
=
\frac{P(s + b)}{P(b)}
=
\frac{
\prob{t \geq \tilde{t}}{s + b}
}{
\prob{t \geq \tilde{t}}{b}\hfill
}
,
\end{equation}
in which the individual \pvalues\ are named
$\clb = P(b)$ and
$\clsb = P(s + b)$.%
\footnote{%
Despite their $\mathrm{CL}$ labels, $\cls$, $\clsb$, and $\clb$ are not
Confidence Levels ---
a confidence level $\gamma$ is a threshold used to assign limits by solving
for $x$ such that $\gamma = 1 - P(x)$~\cite{pdg1996}.
In particle physics, \pvalues\ have sometimes been labelled $\mathrm{CL}$ since
at least 1984~\cite{pdg1984}, although textbooks from the '70s (which are cited
by~\cite{pdg1984}) do distinction the two~\cite{
eadie1971statistical,
frodesen1979probability
}.%
}%
\footnote{%
Some literature defines $\clb = 1 - p_b$~\cite{
Cowan:2010js,
pdg2022ynf
},
in which $p_b$ might not be a \pvalue\ if it excludes equality with the
observed test statistic~\cite{cern2011procedure}.
}

@@

How can one escape this trap?
Likelihood ratios on the actual data escape immediately by bypassing the
censor, but even inequalities need not be biased --- if taken against a
predetermined threshold, say $5$ events, then models can predict whether
$n \leq 5$, and the honestly answered truth or falsehood in data can favour
either side.
Ultimately, a reader may choose to distrust the reporter ---
the messages
`logically, $n \leq 5$' and
`\atlas\ reports $n \leq 5$' and  are not
identical, since one might infer from the latter that in fact the original
data were $n = 5$, before their reduction into a test statistic tail.
The message itself is data, and what it says may not be the whole truth.
